Linked Data Notifications: a resource-centric communication protocol

### 2017-03-17 index.tex
Add Andrei and update acknowledgements

\documentclass[a4paper]{llncs}
\usepackage{ifxetex}
\ifxetex
\usepackage{fontspec}
\defaultfontfeatures{Ligatures=TeX}
\else
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\fi

@@

\usepackage{caption}
\captionsetup{labelfont=bf}

\usepackage{verbatim}

\usepackage[numbers]{natbib}
\usepackage{dcolumn}
\usepackage[export]{adjustbox}
\usepackage{graphicx}
\usepackage[multidot]{grffile}
\usepackage{url}
\usepackage{xspace}
\usepackage{color}
\usepackage{float}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{todonotes}
\usepackage{hyperref}
\usepackage{enumitem}
\setlist[itemize,enumerate]{leftmargin=*}

@@

                Linked Data Notifications: a resource-centric communication protocol

                    \begin{description}
  \item[Authors] \empty Sarven Capadisli\empty 1\empty a\empty Amy Guy\empty 2\empty b\empty Christoph Lange\empty 1\empty c\empty Sören Auer\empty 1\empty d\empty Tim Berners-Lee\empty 3\empty e
    \end{description}

                    \begin{itemize}
  \item 1Enterprise Information Systems Department, \empty University of Bonn, Bonn, DE\item 2School of Informatics, \empty University of Edinburgh, Edinburgh, UK\item 3Decentralized Information Group, CSAIL, \empty MIT, Cambridge, US
    \end{itemize}

                    \begin{itemize}
  \item ✊\empty info@csarven.ca\item 🐦\empty amy@rhiaro.co.uk\item ∫\empty langec@cs.uni-bonn.de\item ⚛\empty auer@cs.uni-bonn.de\item 🕸\empty timbl@w3.org
    \end{itemize}

\par In this article we describe the Linked Data Notifications (LDN) protocol, which is a \empty W3C Candidate Recommendation. Notifications are sent over the Web for a variety of purposes, for example, by social applications. The information contained within a notification is structured arbitrarily, and typically only usable by the application which generated it in the first place. In the spirit of Linked Data, we propose that notifications should be reusable by multiple authorised applications. Through separating the concepts of {\em senders}, {\em receivers} and {\em consumers} of notifications, and leveraging Linked Data principles of shared vocabularies and URIs, LDN provides a building block for decentralised Web applications. This permits end users more freedom to switch between the online tools they use, as well as generating greater value when notifications from different sources can be used in combination. We situate LDN alongside related initiatives, and discuss additional considerations such as security and abuse prevention measures. We evaluate the protocol’s effectiveness by analysing multiple, independent implementations, which pass a suite of formal tests and can be demonstrated interoperating with each other. To experience the described features please open this document in your Web browser under its canonical URI: {\tt \href{http://csarven.ca/linked-data-notifications}{http://csarven.ca/linked-data-notifications}}.

                        \section{Keywords}
  \label{keywords}

                            \begin{itemize}
  \item \empty Communications protocol\item \empty Decentralisation\item \empty Linked Data\item \empty Social web
    \end{itemize}

@@

\par The {\em Linked Data Notifications (LDN)} protocol describes how servers (receivers) can receive messages pushed to them by applications (senders), as well as how other applications (consumers) may retrieve those messages. Any resource can advertise a receiving endpoint (Inbox) for notification messages. Messages are expressed in RDF, and can contain arbitrary data. It is not dependent on a complete implementation of LDP, but comprises an easy-to-implement subset. LDN is a \empty W3C Candidate Recommendation [\empty 4].

                                Overview of Linked Data Notifications

@@

                                LDN Implementations

                                        Implementation
                                        Class\empty *
                                        Description

                                        \begin{description}
  \item[*] Conformance classes: S – sender, C – consumer, R – receiver.\item[a] Implementations by the authors
    \end{description}

\par Source: \empty https://github.com/w3c/ldn/tree/master/implementations

                                        \empty CarbonLDP
                                        R
                                        Data storage platform (LDP)

                                        \empty dokielia
                                        S,C
                                        Clientside editor and annotator

                                        \empty errola
                                        S
                                        Generic message sending client

                                        \empty Fedora Commons
                                        R
                                        Open source repository platform (LDP)

                                        \empty IndieAnndroid
                                        R
                                        Personal blogging platform

                                        \empty Linked Edit Rules
                                        S
                                        Statistical dataset consistency checker

                                        \empty mayktsoa
                                        R
                                        Personal data store (LDP)

                                        \empty OnScreena
                                        C
                                        Notifications display client

                                        \empty pyldn
                                        R
                                        Standalone Inbox

                                        \empty RDF-LinkedData-Notifications
                                        R
                                        Standalone Inbox

                                        \empty slopha
                                        S,R
                                        Social publishing \& quantified self

                                        \empty Solid Words
                                        S
                                        Foreign language learning app

                                        \empty solid-client
                                        S
                                        Clientside library for LDP

                                        \empty solid-inbox
                                        C
                                        Clientside social message reader

                                        \empty solid-notifications
                                        S,C
                                        Clientside library for LDN

                                        \empty solid-server
                                        R
                                        Personal data storage server (LDP)

                                        \empty Virtuoso+\empty ODS Briefcase
                                        R,C
                                        Personal data storage server (LDP)

\par We highlight social scholarly communication use cases with \empty dokieli, a clientside editor for decentralised scientific article publishing, annotations and social interactions [\empty 17]. dokieli uses LDN to send and consume notifications: When a reader comments on a fragment of text in an article, the application discovers the article’s Inbox and sends a notification about the annotation. dokieli also consumes notifications from this Inbox to fetch and display the annotation as marginalia (\empty figure 2). A reader can share a dokieli-enabled article with their contacts; dokieli discovers each contact’s Inbox and sends a notification there (\empty figure 3). When editing an article, the author can add a citation. If an Inbox is discovered in the cited article, dokieli sends a notification there to indicate what part of the article was cited by whom and where. dokieli-enabled articles also consume citation notifications to display these metrics for the author and other readers (\empty figure 4).

                                \empty Video of dokieli Web Annotation

                                \empty Video of dokieli Share

@@

                            \begin{itemize}
  \item If a {\em Like} activity is generated on the server, sloph uses the {\tt object} of the {\em Like} as the target for a notification. Since dokieli uses the same vocabulary for social interactions (\empty ActivityStreams 2.0 [\empty 18]), if the target is a dokieli article, this {\em Like} will be displayed (\empty figure 5).\item If the user publishes a blog post containing a link, which may be semantically annotated to indicate the reason for linking, sloph sends a notification to any Inbox discovered at that link.\item As a receiver, sloph accepts all incoming notifications, but holds for moderation (i.e. places behind access control) any that it cannot automatically verify refer to third-party content published on another domain. If an article written with dokieli publishes a citation of a blog post which advertises a sloph Inbox, sloph will fetch the article and verify whether the relation matches the contents of the notification before exposing the notification for re-use.
    \end{itemize}

\par \empty Linked Edit Rules and \empty Solid Words are specialised senders. Linked Edit Rules checks the consistency of statistical datasets against structured constraints, and delivers the consistency report as a notification to the user. Solid Words is a clientside game for learning new words in a foreign language; it delivers the player’s score for each round to their Inbox. \empty OnScreen is a (crude) generic consumer; as such, it can display notifications sent by both of the aforementioned senders (\empty figure 6).

                                A {\em Like} notification created by sloph, displayed by dokieli.

                                A: Solid Words (a sender), B: Linked Edit Rules (a sender), C: OnScreen (a consumer) displaying notifications sent by A and B.

@@

                                        Comparison of notification mechanisms

                                                Mechanism
                                                T1
                                                T2
                                                T3
                                                R1
                                                R2
                                                R3
                                                R4-A
                                                R4-B
                                                R4-Cp
                                                R4-Cv
                                                R4-Co
                                                R5

                                                    \begin{description}
  \item[T1] Notification type\item[T2] Delivery method\item[T3] Dependencies\item[R1] Modularity (application classes: S Sender, R Receiver, C Consumer, U Subscriber/User)\item[R2] Reusability\item[R3] Persistence - required? how?\item[R4-A] Target representation\item[R4-B] Notification body\item[R4-Cp] Payload processing required?\item[R4-Cv] Verification - required? how?\item[R4-Co] Requirements for referenced resources?\item[R5] Subscription
    \end{description}

                                                    \begin{description}
  \item[–] not applicable, out of scope\item[/] not specified, in scope\item[X] explicitly disallowed\item[app] application specific decision\item[!] required ({\em MUST})\item[+] recommended ({\em SHOULD})\item[O] optional ({\em MAY})\item[PuSH] PubSubHubbub
    \end{description}

                                                    \begin{description}
  \item[h] HTML recommended\item[j] Alternate RDF formats can be negotiated\item[k] {\tt source} and {\tt target} key–value pairs is required\item[q] Provenance records with \empty PROV Ontology\item[r] RDF representation recommended\item[ra] SPARQL results transformed to RSS/Atom\item[s] \empty Sitemaps\item[t] Described in an RDF store or dataset
    \end{description}

                                                Semantic Pingback
                                                Linkback
                                                POST
                                                RDF
                                                S R
                                                /
                                                /
                                                Anyr
                                                form urlencodedk
                                                !
                                                ! parse source
                                                Anyr
                                                X

                                                Webmention
                                                Linkback
                                                POST
                                                HTML
                                                S R
                                                –
                                                –
                                                Anyh
                                                form urlencodedk
                                                !
                                                ! parse source
                                                Anyh
                                                X

                                                Provenance Pingback
                                                Linkback
                                                POST
                                                RDF
                                                S R
                                                /
                                                /
                                                /
                                                URI list
                                                /
                                                /
                                                RDFq
                                                X

                                                DSNotify
                                                Fat ping
                                                POST, PUT
                                                XML, PuSH
                                                S U
                                                /
                                                –
                                                –
                                                XML
                                                /
                                                –
                                                RDFt
                                                !

                                                sparqlPuSH
                                                Fat ping
                                                POST
                                                XML, SPARQL, PuSH
                                                S U
                                                –
                                                –
                                                –
                                                XMLra
                                                /
                                                –
                                                RDFt
                                                !

                                                ResourceSync
                                                Fat ping
                                                POST
                                                XML, PuSH
                                                S U
                                                /
                                                –
                                                –
                                                XMLs
                                                /
                                                –
                                                ?
                                                !

                                                Linked Data Notifications
                                                Fat ping
                                                POST
                                                JSON-LD
                                                S R C
                                                !
                                                ! URI
                                                Any
                                                JSON-LDj
                                                + app
                                                + app
                                                –
                                                O app

@@

                            \begin{enumerate}
  \item Mansour, E., Sambra, A., Hawke, S., Zereba, M., Capadisli, S., Ghanem, A., Aboulnaga, A., Berners-Lee, T.: A Demonstration of the Solid Platform for Social Web Applications, WWW, Demo, 2016, \empty http://www2016.net/proceedings/companion/p223.pdf\item Tramp, S., Frischmuth, P., Ermilov, T., Shekarpour, S., Auer, S.: An Architecture of a Distributed Semantic Social Network, Semantic Web Journal, 2012, \empty \item Arndt, N., Junghanns, K., Meissner, R., Frischmuth, F., Radtke, N., Frommhold, M., Martin, M.: Structured Feedback, WWW, LDOW, 2016, \empty http://events.linkeddata.org/ldow2016/papers/LDOW2016\_paper\_02.pdf\item Capadisli, S., Guy, A.: Linked Data Notifications, W3C Candidate Recommendation, 2016, \empty https://www.w3.org/TR/ldn/\item Parecki, A.: Webmention, W3C Proposed Recommendation, 2016, \empty https://www.w3.org/TR/webmention/\item Langridge, S., Hickson, I.: Pingback 1.0, 2002, \empty http://www.hixie.ch/specs/pingback/pingback\item Klyne, G., Groth, P.: PROV-AQ: Provenance Access and Query,  W3C Note, 2013, \empty http://www.w3.org/TR/prov-aq/\item Haslhofer, B., Popitsch, N.: DSNotify – Detecting and Fixing Broken Links in Linked Data Sets, WWW, 2010, \empty http://eprints.cs.univie.ac.at/81/1/2010\_WWW\_DSNotify.pdf\item Passant, A., Mendes, P.N.: sparqlPuSH: Proactive notification of data updates in RDF stores using PubSubHubbub, SFSW, CEUR Workshop Proceedings, Vol. 699, 2010, \empty http://ceur-ws.org/Vol-699/Paper6.pdf\item Klein, M., Van de Sompel, H., Warner, S., Klyne, G., Haslhofer, B., Nelson, M., Lagoze, C., Sanderson, R.: ResourceSync Framework Specification – Change Notification, 2016, \empty http://www.openarchives.org/rs/notification/1.0/notification\item Berners-Lee, T.: Cool URIs don't change, W3C, 1998, \empty https://www.w3.org/Provider/Style/URI.html\item Archer, P., Loutas, N., Goedertier S., Kourtidis, S.: Study On Persistent URIs, 2012, \empty http://philarcher.org/diary/2013/uripersistence/\item Fielding, R. T.: Architectural Styles and the Design of Network-based Software Architectures. Doctoral dissertation, University of California, Irvine, 2000, \empty http://www.ics.uci.edu/\~fielding/pubs/dissertation/rest\_arch\_style.htm\item Page, K.R., De Roure, D.C., Martinez, K.: REST and Linked Data: a match made for domain driven development?, WWW, WS-REST, 2011, \empty http://ws-rest.org/2011/proc/a5-page.pdf\item Knublauch, H., Kontokostas, D.: Shapes Constraint Language, W3C Working Draft, 2016, \empty \item Speicher, S., Arwe, J., Malhotra, A.: Linked Data Platform, W3C Recommendation, 2015, \empty https://www.w3.org/TR/ldp/\item Capadisli, S., Guy, A., Auer S., Berners-Lee, T.: dokieli, 2016, \empty http://csarven.ca/dokieli\item Snell, J., Prodromou, E.: Activity Streams 2.0, W3C Candidate Recommendation, 2016, \empty https://www.w3.org/TR/activitystreams-core/\item Webber, C., Tallon, J.: ActivityPub, W3C Candidate Recommendation, 2016, \empty https://www.w3.org/TR/activitypub/\item Sporny, M.: JSON-LD and Why I Hate the Semantic Web, 2014, \empty http://manu.sporny.org/2014/json-ld-origins-2/
    \end{enumerate}

>>>>

\documentclass[a4paper]{llncs}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[scaled=.8]{beramono}
\usepackage{latexsym}
\usepackage{textcomp}

@@

\usepackage{graphicx}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{ctable}
\usepackage{array}
\usepackage{hyperref}
\usepackage{enumitem}
\setlist[itemize,enumerate]{leftmargin=*}

@@

\title{Linked Data Notifications: a resource-centric communication protocol}

\author{Sarven Capadisli\inst{1}
  \and Amy Guy\inst{2}
  \and Christoph Lange\inst{1,3}
  \and Sören Auer\inst{1,3}
  \and Andrei Sambra\inst{4}
  \and Tim Berners-Lee\inst{4}}
\institute{%
  University of Bonn, Bonn, DE
  \email{info@csarven.ca, \{langec,auer\}@cs.uni-bonn.de}
  \and School of Informatics, University of Edinburgh, Edinburgh, UK
  \email{amy@rhiaro.co.uk}
  \and Fraunhofer IAIS, Sankt Augustin, DE
  \and Decentralized Information Group, CSAIL, MIT, Cambridge, US
  \email{deiu@mit.edu, timbl@w3.org}
}
\maketitle

\begin{abstract}
  In this article we describe the Linked Data Notifications (LDN) protocol, which is a W3C Candidate Recommendation. Notifications are sent over the Web for a variety of purposes, for example, by social applications. The information contained within a notification is structured arbitrarily, and typically only usable by the application which generated it in the first place. In the spirit of Linked Data, we propose that notifications should be reusable by multiple authorised applications. Through separating the concepts of {\em senders}, {\em receivers} and {\em consumers} of notifications, and leveraging Linked Data principles of shared vocabularies and URIs, LDN provides a building block for decentralised Web applications. This permits end users more freedom to switch between the online tools they use, as well as generating greater value when notifications from different sources can be used in combination. We situate LDN alongside related initiatives, and discuss additional considerations such as security and abuse prevention measures. We evaluate the protocol’s effectiveness by analysing multiple, independent implementations, which pass a suite of formal tests and can be demonstrated interoperating with each other. To experience the described features please open this document in your Web browser under its canonical URI: {\tt \href{http://csarven.ca/linked-data-notifications}{http://csarven.ca/linked-data-notifications}}.
  \keywords{Communications protocol, Decentralisation, Linked Data, Social web}
\end{abstract}

@@

\par The {\em Linked Data Notifications (LDN)} protocol describes how servers (receivers) can receive messages pushed to them by applications (senders), as well as how other applications (consumers) may retrieve those messages. Any resource can advertise a receiving endpoint (Inbox) for notification messages. Messages are expressed in RDF, and can contain arbitrary data. It is not dependent on a complete implementation of LDP, but comprises an easy-to-implement subset.

\begin{figure}
  \centering
  \includegraphics[trim=.2cm 10cm .5cm 10cm,clip,width=.6\textwidth]{media/images/{linked-data-notifications-overview.svg}.pdf}
  \caption{Overview of Linked Data Notifications}
  \label{fig:overview}
\end{figure}

@@

\ctable[
  caption={LDN Implementations},
  label = {tab:implementations}
]{lll}{
  \tnote[*]{Conformance classes: S – sender, C – consumer, R – receiver.}
  \tnote[a]{Implementations by the authors}
  \tnote[]{Source: \url{https://github.com/w3c/ldn}}
}{\FL
  Implementation & Class\tmark[*] & Description
  \ML
                                        CarbonLDP &
                                        R &
                                        Data storage platform (LDP) \NN

                                        dokieli\tmark[a] &
                                        S,C &
                                        Clientside editor and annotator \NN

                                        errol\tmark[a] &
                                        S &
                                        Generic message sending client \NN

                                        Fedora Commons &
                                        R &
                                        Open source repository platform (LDP) \NN

                                        IndieAnndroid &
                                        R &
                                        Personal blogging platform \NN

                                        Linked Edit Rules &
                                        S &
                                        Statistical dataset consistency checker \NN

                                        mayktso\tmark[a] &
                                        R &
                                        Personal data store (LDP) \NN

                                        OnScreen\tmark[a] &
                                        C &
                                        Notifications display client \NN

                                        pyldn &
                                        R &
                                        Standalone Inbox \NN

                                        RDF-LinkedData-Notifications &
                                        R &
                                        Standalone Inbox \NN

                                        sloph\tmark[a] &
                                        S,R &
                                        Social publishing \& quantified self \NN

                                        Solid Words &
                                        S &
                                        Foreign language learning app \NN

                                        solid-client &
                                        S &
                                        Clientside library for LDP \NN

                                        solid-inbox &
                                        C &
                                        Clientside social message reader \NN

                                        solid-notifications &
                                        S,C &
                                        Clientside library for LDN \NN

                                        solid-server &
                                        R &
                                        Personal data storage server (LDP) \NN

                                        Virtuoso+ ODS Briefcase &
                                        R,C &
                                        Personal data storage server (LDP) \LL
}

\par We highlight social scholarly communication use cases with \empty dokieli, a clientside editor for decentralised scientific article publishing, annotations and social interactions~\cite{ref-17}. dokieli uses LDN to send and consume notifications: When a reader comments on a fragment of text in an article, the application discovers the article’s Inbox and sends a notification about the annotation. dokieli also consumes notifications from this Inbox to fetch and display the annotation as marginalia (figure~\ref{fig:video-annotation}). A reader can share a dokieli-enabled article with their contacts; dokieli discovers each contact’s Inbox and sends a notification there (figure~\ref{fig:video-share}). When editing an article, the author can add a citation. If an Inbox is discovered in the cited article, dokieli sends a notification there to indicate what part of the article was cited by whom and where. dokieli-enabled articles also consume citation notifications to display these metrics for the author and other readers (figure~\ref{fig:like}).

\begin{figure}
  \begin{minipage}[b]{.49\textwidth}
    \includegraphics[width=\textwidth]{media/images/dokieli-annotation}
    \caption{Video of dokieli Web Annotation}
    \label{fig:video-annotation}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{.49\textwidth}
    \includegraphics[width=\textwidth]{media/images/dokieli-share}
    \caption{Video of dokieli Share}
    \label{fig:video-share}
  \end{minipage}
\end{figure}

@@

                            \begin{itemize}
  \item If a {\em Like} activity is generated on the server, sloph uses the {\tt object} of the {\em Like} as the target for a notification. Since dokieli uses the same vocabulary for social interactions (\empty ActivityStreams 2.0~\cite{ref-18}), if the target is a dokieli article, this {\em Like} will be displayed (figure~\ref{fig:like}).\item If the user publishes a blog post containing a link, which may be semantically annotated to indicate the reason for linking, sloph sends a notification to any Inbox discovered at that link.\item As a receiver, sloph accepts all incoming notifications, but holds for moderation (i.e. places behind access control) any that it cannot automatically verify refer to third-party content published on another domain. If an article written with dokieli publishes a citation of a blog post which advertises a sloph Inbox, sloph will fetch the article and verify whether the relation matches the contents of the notification before exposing the notification for re-use.
    \end{itemize}

\par \empty Linked Edit Rules and \empty Solid Words are specialised senders. Linked Edit Rules checks the consistency of statistical datasets against structured constraints, and delivers the consistency report as a notification to the user. Solid Words is a clientside game for learning new words in a foreign language; it delivers the player’s score for each round to their Inbox. \empty OnScreen is a (crude) generic consumer; as such, it can display notifications sent by both of the aforementioned senders (figure~\ref{fig:notifications}).

\begin{figure}
  \includegraphics[width=\textwidth]{media/images/screenshot-ldn-sloph-dokieli}
  \caption{A {\em Like} notification created by sloph, displayed by dokieli.}
  \label{fig:like}

  \includegraphics[width=\textwidth]{media/images/screenshot-ldn-senders}
  \caption{A: Solid Words (a sender), B: Linked Edit Rules (a sender), C: OnScreen (a consumer) displaying notifications sent by A and B.}
  \label{fig:notifications}
\end{figure}

@@

\begin{table}
  \caption{Comparison of notification mechanisms}
  \label{tab:comparison}
  \footnotesize
  \begin{tabular}{% Mechanism
    p{.14\textwidth}
    % T1–T3
    p{.065\textwidth}
    p{.075\textwidth}
    p{.1\textwidth}
    % R1–R3
    p{.03\textwidth}
    p{.03\textwidth}
    p{.04\textwidth}
    % R4-A–R5
    p{.06\textwidth}
    p{.14\textwidth}
    p{.045\textwidth}
    p{.06\textwidth}
    p{.06\textwidth}
    p{.05\textwidth}
    }\FL
                                                Mechanism&
                                                T1&
                                                T2&
                                                T3&
                                                R1&
                                                R2&
                                                R3&
                                                R4-A&
                                                R4-B&
                                                R4-C\textsuperscript{p}&
                                                R4-C\textsuperscript{v}&
                                                R4-C\textsuperscript{o}&
                                                R5\ML
                                                Semantic\newline Pingback&
                                                Link-\newline back&
                                                POST&
                                                RDF&
                                                S\newline R&
                                                /&
                                                /&
                                                Any\textsuperscript{r}&
                                                form\newline urlencoded\textsuperscript{k}&
                                                !&
                                                !\newline parse\newline source&
                                                Any\textsuperscript{r}&
                                                X\NN
                                                Web-\newline mention&
                                                Link-\newline back&
                                                POST&
                                                HTML&
                                                S\newline R&
                                                –&
                                                –&
                                                Any\textsuperscript{h}&
                                                form\newline urlencoded\textsuperscript{k}&
                                                !&
                                                !\newline parse\newline source&
                                                Any\textsuperscript{h}&
                                                X\NN
                                                Provenance\newline Pingback&
                                                Link-\newline back&
                                                POST&
                                                RDF&
                                                S\newline R&
                                                /&
                                                /&
                                                /&
                                                URI list&
                                                /&
                                                /&
                                                RDF\textsuperscript{q}&
                                                X\NN
                                                DSNotify&
                                                Fat\newline ping&
                                                POST,\newline PUT&
                                                XML,\newline PuSH&
                                                S\newline U&
                                                /&
                                                –&
                                                –&
                                                XML&
                                                /&
                                                –&
                                                RDF\textsuperscript{t}&
                                                !\NN
                                                sparqlPuSH&
                                                Fat\newline ping&
                                                POST&
                                                XML,\newline SPARQL,\newline PuSH&
                                                S\newline U&
                                                –&
                                                –&
                                                –&
                                                XML\textsuperscript{ra}&
                                                /&
                                                –&
                                                RDF\textsuperscript{t}&
                                                !\NN
                                                Resource-\newline Sync&
                                                Fat\newline ping&
                                                POST&
                                                XML,\newline PuSH&
                                                S\newline U&
                                                /&
                                                –&
                                                –&
                                                XML\textsuperscript{s}&
                                                /&
                                                –&
                                                ?&
                                                !\NN
                                                Linked Data\newline Notifications&
                                                Fat\newline ping&
                                                POST&
                                                JSON-LD&
                                                S\newline R\newline C&
                                                !&
                                                !\newline URI&
                                                Any&
                                                JSON-LD\textsuperscript{j}&
                                                + app&
                                                + app&
                                                –&
                                                O\newline app\LL
  \end{tabular}\\
  \begin{tabularx}{1.0\linewidth}{XX}\FL
    \textbf{T1}: Notification type&
    \textbf{R4-A}: Target representation\NN
    \textbf{T2}: Delivery method&
    \textbf{R4-B}: Notification body\NN
    \textbf{T3}: Dependencies&
    \textbf{R4-C\textsuperscript{p}}: Payload processing required?\NN
    \textbf{R1}: Modularity (application classes: S Sender, R Receiver, C Consumer, U Subscriber/User)&
    \textbf{R4-C\textsuperscript{v}}: Verification - required? how?\NN
    \textbf{R2}: Reusability&
    \textbf{R4-C\textsuperscript{o}}: Requirements for referenced resources?\NN
    \textbf{R3}: Persistence - required? how?&
    \textbf{R5}: Subscription\ML
    \textbf{–}: not applicable, out of scope&
    \textbf{!}: required ({\em MUST})\NN
    \textbf{/}: not specified, in scope&
    \textbf{+}: recommended ({\em SHOULD})\NN
    \textbf{X}: explicitly disallowed&
    \textbf{O}: optional ({\em MAY})\NN
    \textbf{app}: application specific decision&
    \textbf{PuSH}: PubSubHubbub\ML
    \textbf{\textsuperscript{h}}: HTML recommended&
    \textbf{\textsuperscript{r}}: RDF representation recommended\NN
    \textbf{\textsuperscript{j}}: Alternate RDF formats can be negotiated&
    \textbf{\textsuperscript{ra}}: SPARQL results transformed to RSS/Atom\NN
    \textbf{\textsuperscript{k}}: {\tt source} and {\tt target} key–value pairs is required&
    \textbf{\textsuperscript{s}}: \empty Sitemaps\NN
    \textbf{\textsuperscript{q}}: Provenance records with \empty PROV Ontology&
    \textbf{\textsuperscript{t}}: Described in an RDF store or dataset\LL
  \end{tabularx}
\end{table}

@@

\paragraph{Acknowledgements:}
This material is based on work supported by the Qatar Computing Research Institute (QCRI), and by the DFG project “Opening Scholarly Communication in Social Sciences” (grant agreement AU 340/9-1).

\begin{thebibliography}{99}
  \bibitem{ref-1} Mansour, E., Sambra, A., Hawke, S., Zereba, M., Capadisli, S., Ghanem, A., Aboulnaga, A., Berners-Lee, T.: A Demonstration of the Solid Platform for Social Web Applications, WWW, Demo, 2016%, \url{http://www2016.net/proceedings/companion/p223.pdf}
  \bibitem{ref-2} Tramp, S., Frischmuth, P., Ermilov, T., Shekarpour, S., Auer, S.: An Architecture of a Distributed Semantic Social Network, Semantic Web Journal, 2012
  \bibitem{ref-3} Arndt, N., Junghanns, K., Meissner, R., Frischmuth, F., Radtke, N., Frommhold, M., Martin, M.: Structured Feedback, WWW, LDOW, 2016%, \url{http://events.linkeddata.org/ldow2016/papers/LDOW2016_paper_02.pdf}
  \bibitem{ref-4} Capadisli, S., Guy, A.: Linked Data Notifications, W3C Candidate Recommendation, 2016, \url{https://www.w3.org/TR/ldn/}
  \bibitem{ref-5} Parecki, A.: Webmention, W3C Proposed Recommendation, 2016,  \url{https://www.w3.org/TR/webmention/}
  \bibitem{ref-6} Langridge, S., Hickson, I.: Pingback 1.0, 2002,  \url{http://www.hixie.ch/specs/pingback/pingback}
  \bibitem{ref-7} Klyne, G., Groth, P.: PROV-AQ: Provenance Access and Query,  W3C Note, 2013, \url{http://www.w3.org/TR/prov-aq/}
  \bibitem{ref-8} Haslhofer, B., Popitsch, N.: DSNotify – Detecting and Fixing Broken Links in Linked Data Sets, WWW, 2010%, \url{http://eprints.cs.univie.ac.at/81/1/2010\_WWW\_DSNotify.pdf}
  \bibitem{ref-9} Passant, A., Mendes, P.N.: sparqlPuSH: Proactive notification of data updates in RDF stores using PubSubHubbub, SFSW, CEUR Workshop Proceedings, Vol. 699, 2010%, \url{http://ceur-ws.org/Vol-699/Paper6.pdf}
  \bibitem{ref-10} Klein, M., Van de Sompel, H., Warner, S., Klyne, G., Haslhofer, B., Nelson, M., Lagoze, C., Sanderson, R.: ResourceSync Framework Specification – Change Notification, 2016,  \url{http://www.openarchives.org/rs/notification/1.0/notification}
  \bibitem{ref-11} Berners-Lee, T.: Cool URIs don't change, W3C, 1998,  \url{https://www.w3.org/Provider/Style/URI.html}
  \bibitem{ref-12} Archer, P., Loutas, N., Goedertier S., Kourtidis, S.: Study On Persistent URIs, 2012, \url{http://philarcher.org/diary/2013/uripersistence/}
  \bibitem{ref-13} Fielding, R. T.: Architectural Styles and the Design of Network-based Software Architectures. Doctoral dissertation, University of California, Irvine, 2000%,  \url{http://www.ics.uci.edu/~fielding/pubs/dissertation/rest_arch_style.htm}
  \bibitem{ref-14} Page, K.R., De Roure, D.C., Martinez, K.: REST and Linked Data: a match made for domain driven development?, WWW, WS-REST, 2011%,  \url{http://ws-rest.org/2011/proc/a5-page.pdf}
  \bibitem{ref-15} Knublauch, H., Kontokostas, D.: Shapes Constraint Language, W3C Working Draft, 2016
  \bibitem{ref-16} Speicher, S., Arwe, J., Malhotra, A.: Linked Data Platform, W3C Recommendation, 2015, \url{https://www.w3.org/TR/ldp/}
  \bibitem{ref-17} Capadisli, S., Guy, A., Auer S., Berners-Lee, T.: dokieli, 2016, \url{http://csarven.ca/dokieli}
  \bibitem{ref-18} Snell, J., Prodromou, E.: Activity Streams 2.0, W3C Candidate Recommendation, 2016, \url{https://www.w3.org/TR/activitystreams-core/}
  \bibitem{ref-19} Webber, C., Tallon, J.: ActivityPub, W3C Candidate Recommendation, 2016, \url{https://www.w3.org/TR/activitypub/}
  \bibitem{ref-20} Sporny, M.: JSON-LD and Why I Hate the Semantic Web, 2014, \url{http://manu.sporny.org/2014/json-ld-origins-2/}
\end{thebibliography}
